Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2020-05-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
add abstract

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
	The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peer-review papers.
\begin{IEEEkeywords}
	hyperspectral imagery, forest health monitoring, machine learning, feature selection, model comparison, imaging spectroscopy
\end{IEEEkeywords}

@@

\begin{figure} [t!]
	\centering
	\begin{center}
		\includegraphics[width=0.48\textwidth] {defoliation-distribution-plot-1.pdf}
		\caption{Response variable \textit{defoliation} at trees for plots \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}. \texttt{n} corresponds to the total number of trees in the plot, $\bar{x}$ refers to the mean defoliation.}\label{fig:defol-distr}
	\end{center}
\end{figure}

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
	Analyzing highly-correlated feature-rich datasets from hyperspectral remote sensing data is a challenging task.
	This study compared multiple machine-learning and statistical learning methods using filter-based feature-selection methods on a hyperspectral dataset.
	In the case study, defoliation at trees (\%) in northern Spain was modeled as a function of band reflectance.
	Multiple feature sets were compared using derived features (expert-based and data-driven).
	Among the selected methods, support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by about three percentage points.
	The combination of feature sets showed small increases in predictive performance while no substantial differences between single feature sets were observed.
	Filter methods helped in certain task-learner-filter settings to achieve better predictive performances while ensemble filters did not have a substantial impact on performance.
	Depending on the dataset-learner combination, filter methods can help to improve predictive performance and lower computational runtime.
	Permutation-based feature importance estimated features around the red edge to be most important for the models.
	However, features in the near-infrared region (800 nm - 1000 nm) were also important to achieve acceptable performances.
	The dataset used in this work was suboptimal and more training data is needed for more generalizable conclusions.
\end{abstract}

% Note that keywords are not normally used for peer-review papers.
\begin{IEEEkeywords}
	hyperspectral imagery, forest health monitoring, machine learning, feature selection, model comparison, filter, imaging spectroscopy
\end{IEEEkeywords}

@@

\begin{figure} [t!]
	\centering
	\begin{center}
		\includegraphics[width=0.48\textwidth] {defoliation-distribution-plot-1.pdf}
		\caption{Response variable \textit{defoliation} at trees for plots Laukiz1, Laukiz2, Luiando and Oiartzun. \texttt{n} corresponds to the total number of trees in the plot, $\bar{x}$ refers to the mean defoliation.}\label{fig:defol-distr}
	\end{center}
\end{figure}
